---
title: "Does Town Size Influence Youth Employment Opportunities?"
author: "Arun Krishnasamy"
date: "2024-10-6"
categories: [Post]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      eval = TRUE, 
                      cache = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
options(width=80, digits=3)
```

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(readxl)
library(janitor)
library(broom)
```

The common perception that moving to a large city automatically leads to better job opportunities has been widely accepted for years. However, this analysis seeks to question that assumption by exploring whether 19-year-olds have better employment opportunities in smaller towns compared to bigger cities. Understanding this dynamic could offer valuable insights for young adults in the early stages of their career planning.

The data used in this analysis comes from the [TidyTuesdayR](https://github.com/rfordatascience/tidytuesday) project, sourced from [The UK office for National Statistics](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/educationandchildcare/datasets/educationalattainmentofyoungpeopleinenglishtownsdata/200708201819/youngpeoplesattainmentintownsreferencetable1.xlsx). It focuses on employment rates for 19-year-olds across different UK towns, categorized as:

- **Employment Rate**: The percentage of 19-year-olds employed with earnings above zero.
- **Size Classification**: Categorized into “Small,” “Medium,” and “Large” towns.

```{r, echo=TRUE}
df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-01-23/english_education.csv')
```

```{r}
df <- read_excel("data/data.xlsx", sheet = 2)
```
Before conducting the analysis, the data went through a cleaning processes. Non-numeric employment data were removed, and town size categories were recoded, excluding outliers such as "Inner London BUA" to ensure consistent comparisons.

```{r, echo=TRUE}
df_clean <- df |>
  clean_names() |>
  filter(activity_at_age_19_employment_with_earnings_above_0 != "*") |>
  mutate(Employment_Rate = as.numeric(activity_at_age_19_employment_with_earnings_above_0), 
         Size_Classification = as.factor(size_flag)) |>
  filter(!Size_Classification %in% c("Inner London BUA", "Outer london BUA", "Not BUA", "Other Small BUAs"))
```
The box plot below (@fig-boxplot) illustrates the distribution of employment rates across different town sizes. Small towns demonstrate a higher median employment rate for 19-year-olds compared to cities, which show both the lowest median and the largest variation in employment outcomes.

```{r}
#| label: fig-boxplot
#| fig-cap: "The distribution of employment rates at age 19 across different town sizes. Cities have the lowest median and largest variation, while Small Towns have the highest variation but a median similar to Medium and Large Towns."
ggplot(df_clean, aes(x = Size_Classification, y = Employment_Rate)) +
  geom_boxplot(fill = "steelblue") +
  labs(title = "Employment Rate Distribution at Age 19 by Town Size",
       x = "Town Size Classification",
       y = "Employment Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

To quantify the employment opportunities in each town size, the table below (@tbl-employment-rate) summarizes the average employment rates for each category. Small towns lead with an average employment rate of 50.1%, while cities fall behind with 42.6%.

```{r}
#| label: tbl-employment-rate
#| tbl-cap: "The average employment rates for 19-year-olds across different town sizes. Small towns exhibit the highest average employment rates, while cities show the lowest."
employment_summary <- df_clean |>
  group_by(Size_Classification) |>
  summarise(Average_Employment_Rate = mean(Employment_Rate, na.rm = TRUE))

kable(employment_summary) |> 
  kable_styling(font_size = 15)
```

To confirm whether these differences in employment rates across town sizes are statistically significant, an ANOVA test was conducted. The results below (@tbl-anova) indicate that there is a statistically significant difference between the town size categories.

```{r}
#| label: tbl-anova
#| tbl-cap: "ANOVA Results for the Effect of Town Size Classification on Employment Rates at Age 19"
anova_result <- aov(Employment_Rate ~ Size_Classification, data = df_clean)
kable(tidy(anova_result)) |> 
  kable_styling(font_size = 10)
```

While ANOVA identifies that differences exist between town sizes, it does not specify which pairs of town sizes differ significantly. The Tukey HSD post-hoc test (see table @tbl-posthoc) addresses this by providing pairwise comparisons between town sizes.

```{r, echo=FALSE}
#| label: tbl-posthoc
#| tbl-cap: "Tukey HSD Post-hoc Test results showing pairwise comparisons between town sizes. Small and Medium Towns exhibit significantly higher employment rates than Cities and Large Towns."
posthoc_result <- TukeyHSD(anova_result) |> 
  tidy() |> 
  select(-term, -null.value)

kable(posthoc_result) |> 
  kable_styling(font_size = 10) 
```

- **Medium Towns vs. City:** Medium towns exhibit significantly higher employment rates than cities.

- **Small Towns vs. City:** Small towns show a notable advantage over cities.

- **Small Towns vs. Large Towns:** Small towns have significantly higher employment rates than large towns.

- **Medium Towns vs. Large Towns:** Medium towns also outperform large towns in employment rates.

- **Small Towns vs. Medium Towns:** A smaller but significant difference was observed between small and medium towns.


The comparison between **Large Towns vs. City** did not show a statistically significant difference (p = 0.107), suggesting relatively similar employment rates between these two categories.

The Tukey HSD results are further visualized in the plot below (@fig-HSD), highlighting the confidence intervals for each comparison. Comparisons where the confidence intervals do not cross zero indicate statistically significant differences.

```{r}
#| label: fig-HSD
#| fig.cap: "The figure provides a quick visual overview of which pairwise comparisons are significant."
tukey_hsd <- ggplot(posthoc_result, aes(x = estimate, y = reorder(contrast, estimate))) +
  geom_point(aes(color = adj.p.value < 0.05), size = 3) + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) + 
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Mean Difference (Estimate)", y = "Contrast", title = "Tukey HSD Post-hoc Test Results") +
  theme_minimal() +
  scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black")) +
  theme(legend.position = "none")
tukey_hsd
```

This analysis provides evidence that smaller towns offer better employment opportunities for young adults compared to larger cities and towns. With small and medium towns showing consistently higher employment rates, it challenges the conventional wisdom that large cities are always the best option for starting a career.

The findings suggest that young job seekers may find more opportunities in smaller towns, which could influence career planning strategies and policy decisions aimed at improving youth employment.

ONS. (2023). Why Do Children and Young People in Smaller Towns Do Better Academically Than Those in Larger Towns? [Office for National Statistics](https://www.ons.gov.uk/peoplepopulationandcommunity/educationandchildcare/articles/whydochildrenandyoungpeopleinsmallertownsdobetteracademicallythanthoseinlargertowns/2023-07-25).

R for Data Science Online Learning Community, 2024. TidyTuesday: A weekly data project aimed at the R ecosystem. GitHub repository<https://github.com/rfordatascience/tidytuesday>

<footer style="text-align: center; padding: 1rem; background-color: #333; color: white;">
  <p>© 2024 Arun Krishnasamy. All rights reserved.</p>
  <ul style="list-style: none; padding: 0;">
    <li><a href="https://github.com/ArunPrakash2901" style="color: white;">GitHub</a></li>
    <li><a href="https://linkedin.com/in/apkrishnasamy" style="color: white;">LinkedIn</a></li>
  </ul>
</footer>
